\sffamily An EEG-based visual Brain-Computer Interface for gaze-free communication

### 2024-11-27 chapters/stbf_struct.tex
Layout accuracy tables

	\section{Materials \& methods}
	\subsection{Notation}
	We represent matrices with bold capital letters, vectors with bold
	lowercase letters, fixed scalars with uppercase cursive letters and variable
  scalars with cursive lowercase letters.
	The epoched \ac{eeg} data with $N$ epochs, $C$ channels, and $S$ samples are
  represented in epoch format as $\{\mat{X}_n\in\mathbb{R}^{C\times S}\}^N_{n=1}$
  or flattened vector format by concatenating all channels for each epoch.
	Flattening results in $\{\mat{x}_n\in\mathbb{R}^{CS}\}^N_{n=1}$ such that
  $\mat{x}_n = \text{vec}(\mat{X}_n)$.
	The real covariance matrix of the epochs in vector format is
  denoted by $\mat{C}$, estimators thereof as $\hat{\mat{C}}$.

	\subsection{Spatiotemporal beamforming}
  \ac{lcmv}-beamforming was initially introduced to \ac{eeg} analysis as a filter for
	source localization~\cite{VanVeen1997} to enhance the signal-to-noise ratio
	(SNR).
	\textcite{VanVliet2015} first applied the spatiotemporal
	\ac{lcmv}-beamformer as a method for the analysis of \acp{erp}.
	The extension to the combined spatiotemporal domain~\cite{VanVliet2015} and the
	data-driven approaches proposed by \textcite{Treder2016} and
	\textcite{Wittevrongel2016} allow for its application to classification problems.

	For the following analyses, we assume that all \ac{eeg} channels are normalized with zero mean and unit variance without loss of generality.
	Solving \cref{eq:stbf-struct/minimum-variance} under the linear constraint given by
  \cref{eq:stbf-struct/linear-constraint} returns the filter weights $\mat{w}$ defining the spatiotemporal \ac{lcmv}-beamformer.
	\begin{equation}
    \argmin_\mat{w}\mat{w}^\intercal \mat{C}
		\mat{w}^\intercal
		\label{eq:stbf-struct/minimum-variance}
	\end{equation}
	\begin{equation}
		\mat{a}^\intercal\mat{w} = 1
		\label{eq:stbf-struct/linear-constraint}
	\end{equation}
	These weights minimize the variance of the output of the filter while enhancing
	the signal characterized by the constraint.
  $\mat{a} = \text{vec}(\mat{A})$ is the data-driven activation pattern, a template
	of the signal of interest maximizing the difference between two classes of
	epochs, determined as follows:
	\begin{equation}
		\mathbf{a} =
		\frac{1}{|\text{class 1}|}\sum_\text{class 1}\mathbf{x_n} -
		\frac{1}{|\text{class 2}|}\sum_\text{class 2}\mathbf{x_n}
		\label{eq:stbf-struct/ap}
	\end{equation}

@@

	\subsubsection{Shrunk covariance estimation}
	\label{sec:stbf-struct/methods/shrunk-cov}
	The shrinkage covariance estimator creates a better conditioned inversion matrix problem and generally performs better when applied to unseen data.
	The estimators for the covariance and inverse covariance are given by:
	\begin{equation}
    \hat{\mat{C}}_\alpha =
    (1-\alpha) \hat{\mat{C}}_\text{emp}
    + \alpha\frac{\text{Tr}(\hat{\mat{C}}_\text{emp})}{CS}\mathbb{I}
		\label{eq:shrinkage}
	\end{equation}
	\begin{equation}
    \widehat{\mat{C}^{-1}}_\alpha =
    \hat{\mat{C}}^+_\alpha
		\label{eq:stbf-struct/shrinkage-inv}
	\end{equation}
	with $0<\alpha<1$.
	Analogous to $L_2$ regularization of the beamforming problem,
	shrinkage reduces the ratio between the smallest and largest eigenvalues
	of the covariance matrix by strengthening the diagonal.
	\cref{fig:kronlda-covs}c and \cref{fig:kronlda-covs}d respectively show examples of the
	shrunk estimator of the covariance and the inverse covariance matrices.

@@

	Methods exist to estimate an optimal shrinkage value from the data directly.
	Most notable among these are the Ledoit-Wolf procedure~\cite{Ledoit2004},
	Rao-Blackwell Ledoit-Wolf~\cite{Chen2010}, and Oracle Approximating Shrinkage~\cite{Chen2010}.
  A more recent estimation method~\cite{Tong2018} emulates a \ac{loocv} scheme expressed by the data-driven closed-form
	estimate:
	\begin{equation}
		\alpha =
		1-\frac{
      \frac{N}{N-1}\text{Tr}(\hat{\mat{C}}_\text{emp}^2)
      - \frac{2}{CS}\left[\text{Tr}(\hat{\mat{C}}_\text{emp})\right]^2
      + \frac{1}{CS}\text{Tr}(\hat{\mat{C}}_\text{emp}^2)
			- \frac{1}{N(N-1)}\sum_{n=1}^N||\mat{x}_n||_2^4
		}
		{
      \frac{N^2 -2N}{(N-1)^2}\text{Tr}(\hat{\mat{C}}_\text{emp}^2)
      - \frac{2}{CS}\left[\text{Tr}(\hat{\mat{C}}_\text{emp})\right]^2
      + \frac{1}{CS}\text{Tr}(\hat{\mat{C}}_\text{emp}^2)
			+ \frac{1}{N(N-1)^2}\sum_{n=1}^N||\mat{x}_n||_2^4
		}
		\label{eq:loocv}
	\end{equation}
	We opt for the \ac{loocv} shrinkage estimator because it avoids some of the
	assumptions made by \textcite{Ledoit2004} and \textcite{Chen2010} and
	because it generalizes to structured covariance estimation as described in
	\cref{sec:stbf-struct/methods/structured-estimation}.

@@

	Combined, this leads to the iterative estimation algorithm described by the
	following equations:
	\begin{subequations}
		\begin{equation}
      \tilde{\mat{S}}_{k+1} =
			\frac{1}{N}
      \sum^N_{n=1}\mat{X}_n^\intercal\hat{\mat{T}}_k^+\mat{X}_n
			\label{eq:stbf-struct/fpi-spatial}
		\end{equation}
		\begin{equation}
      \tilde{\mat{T}}_{k+1} =
			\frac{1}{N}
      \sum^N_{n=1}\mat{X}_n\hat{\mat{S}}_k^+\mat{X}_n^\intercal
			\label{eq:stbf-struct/fpi-temporal}
		\end{equation}
	\end{subequations}
	\begin{subequations}
		\begin{equation}
      \tilde{\mat{S}}_{k+1}^{(\beta)} =
      (1-\beta_{k+1})\tilde{\mat{S}}_{k+1}
      +\beta_{k+1}\frac{\text{Tr}(\tilde{\mat{S}}_{k+1})}{C}\mathbb{I}
			\label{eq:stbf-struct/fpi-spatial-shrunk}
		\end{equation}
		\begin{equation}
      \tilde{\mat{T}}_{k+1}^{(\gamma)} =
      (1-\gamma_{k+1})\tilde{\mat{T}}_{k+1}
      +\gamma_{k+1}\frac{\text{Tr}(\tilde{\mat{T}}_{k+1})}{S}\mathbb{I}
			\label{eq:stbf-struct/fpi-temporal-shrunk}
		\end{equation}
	\end{subequations}
	\begin{subequations}
		\begin{equation}
      \hat{\mat{S}}_{k+1} =
      \frac{C}{\text{Tr}\left[\tilde{\mat{S}}_{k+1}^{(\beta)}\right]}
      \tilde{\mat{S}}_{k+1}^{(\beta)}
			\label{eq:stbf-struct/fpi-spatial-norm}
		\end{equation}
		\begin{equation}
      \hat{\mat{T}}_{k+1} =
      \frac{S}{\text{Tr}\left[\tilde{\mat{T}}_{k+1}^{(\gamma)}\right]}
      \tilde{\mat{T}}_{k+1}^{(\gamma)}
			\label{eq:stbf-struct/fpi-temporal-norm}
		\end{equation}
	\end{subequations}
  $\hat{\mat{S}}_0$ and $\hat{\mat{T}}_0$ can be initialized to any positive definite matrix.
	We choose to use the identity matrices $\mathbb{I}^{C\times C}$ and $\mathbb{I}^{S\times S}$.
  After each iteration, all diagonals of $\hat{\mat{R}}_{k+1}$ are set to their mean
  values to ensure that $\hat{\mat{R}}_{k+1}$ and $\hat{\mat{T}}_{k+1}$ are Toeplitz structured.

@@

	\begin{subequations}
		\begin{equation}
			\beta_{k+1} =
			1-
			\frac{
        \frac{N}{N-1}\text{Tr}(\tilde{\mat{\hat{\S}}}_{k+1}^2)
        - \frac{2}{C}\left[\text{Tr}(\mat{\tilde{S}}_{k+1})\right]^2
        + \frac{1}{C}\text{Tr}(\tilde{\mat{S}}_{k+1}^2)
				- \frac{1}{N(N-1)}\sum_{i=1}^N
        \left[\text{Tr}(\mat{X}_i\hat{\mat{T}}_k^+\mat{X}_i^\intercal)^2\right]
			}{
        \frac{N^2-2N}{(N-1)^2}\text{Tr}(\tilde{\mat{S}}_{k+1}^2)
        - \frac{2}{C}\left[\text{Tr}(\tilde{\mat{S}}_{k+1})\right]^2
        + \frac{1}{C}\text{Tr}(\tilde{\mat{S}}_{k+1}^2)
				+ \frac{1}{N(N-1)^2}\sum_{i=1}^N
        \left[\text{Tr}(\mat{X}_i\hat{\mat{T}}_k^+\mat{X}_i^\intercal)^2\right]
			}
			\label{eq:stbf-struct/spatial-shrinkage}
		\end{equation}
		\begin{equation}
			\gamma_{k+1} =
			1-
			\frac{
        \frac{N}{N-1}\text{Tr}(\tilde{\mat{T}}_{k+1}^2)
        - \frac{2}{S}\left[\text{Tr}(\tilde{\mat{T}}_{k+1})\right]^2
        + \frac{1}{S}\text{Tr}(\tilde{\mat{T}}_{k+1}^2)
				- \frac{1}{N(N-1)}\sum_{n=1}^N
        \left[\text{Tr}(\mat{X}_n^\intercal\mat{\mat{S}}_k^+\mat{X}_n)^2\right]
			}{
        \frac{N^2-2N}{(N-1)^2}\text{Tr}(\tilde{\mat{T}}_{k+1}^2)
        - \frac{2}{S}\left[\text{Tr}(\tilde{\mat{T}}_{k+1})\right]^2
        + \frac{1}{S}\text{Tr}(\tilde{\mat{T}}_{k+1}^2)
				+ \frac{1}{N(N-1)^2}\sum_{n=1}^N
        \left[\text{Tr}(\mat{X}_n^\intercal\hat{\mat{S}}_k^+\mat{X}_n)^2\right]
			}
			\label{eq:stbf-struct/temporal-shrinkage}
		\end{equation}
	\end{subequations}
	The shrinkage parameters $0<\beta_{k+1}<1$ and $0<\gamma_{k+1}<1$ should be
	re-determined after each iteration.
	The Oracle Approximation Shrinkage method can also be used to determine
	$\beta_{k+1}$ and $\gamma_{k+1}$~\cite{Chen2010,Xie2021} but performs worse for spatiotemporal \ac{eeg} data since not all assumptions are met.

@@

  \subsubsection{\Acl{stbf} classifier}
	Before calculating the \ac{stbf}, the signal was downsampled to
	32 Hz or twice the low-pass frequency 16 Hz, resulting in 17 time samples
	between 0.1 and 0.6s. According to the Nyquist Theorem, more samples would not
	contain more information hence the minimum temporal resolution is chosen to reduce
	the dimensionality of the covariance and improve its condition number.
	The activation pattern is the difference between the averages of epochs in response to cued targets and the averages of those in response to non-cued targets.
	We constructed three variations of the spatiotemporal beamformer:
	\ac{stbf} with empirical covariance estimation (\textsc{stbf-emp}) as in
	\cref{sec:stbf-struct/methods/emp-cov}, \ac{stbf} with
	\ac{loocv}-shrunk covariance estimation (\textsc{stbf-shrunk}) as in
	\cref{sec:stbf-struct/methods/shrunk-cov}, and \ac{stbf} with
	Kronecker-Toeplitz structured covariance estimation (\textsc{stbf-struct}) with \ac{loocv} shrinkage for
	the Kronecker factors as in \cref{sec:stbf-struct/methods/struct-cov}.

	\subsubsection{Riemannian geometry classifier}
	\label{sec:riemannian}
	We opted for a Riemannian
	geometry-based classifier to compare our results.
	The Riemannian model (\textsc{xdawn+rg}) uses the XDAWN spatial filter combined
	with Riemannian geometry in tangent space as implemented by Barachant et
	al.~\cite{Barachant2014a}.
	This classifier uses four XDAWN spatial filters and each epoch's empirical spatial covariance matrix.
	The target with the maximum score is the prediction of the cued target.
	\textsc{xdawn+rg} was trained and validated without downsampling using epochs
	at the original sample rate of 2048 Hz.
  Preliminary experimentation revealed that best performance in low and high
  data availability setting was achieved when no shrinkage was applied to the estimation of XDAWN filters or the subsequent
  covariance matrices for this specific dataset.
  Hence, no shrinkage regularization will be applied to \textsc{xdawn+rg} in
  further experiments.

	\section{Results}
	\subsection{Minimum required fixed-point iterations}
	The fixed point iteration algorithm described
	in
  \crefrange{eq:stbf-struct/fpi-spatial}{eq:stbf-struct/fpi-temporal-norm}
  estimates the Kronecker-Toeplitz structured covariance for the
	\textsc{stbf-struct} classifier.
	Fixed-point iteration is an iterative procedure starting from (in our case)
	non-informed initial guesses for the spatial and temporal covariance matrices.
	As a stopping criterion, one could impose a threshold on the difference in
	outcome of successive steps, e.g., based on the covariance norm or the
	classifier accuracy.
	However, few iterations or even just one~\cite{Castaneda2014} suffice to achieve satisfactory performance in practice.

	\cref{fig:iterations} confirms these results for the \textsc{stbf-struct} classifier.
	Using more than one fixed-point iteration does not significantly improve the
	accuracy across the amounts of training data and the number of trials
  used for evaluation.
	Hence, only one iteration is used for the \textsc{stbf-struct} classifier, leading to a drastic speed-up of
  the  training process.

  \begin{figure}
    \sffamily
    \begin{subfigure}{\linewidth}
      \input{figures/stbf_struct/fpi-0.pgf}
      \caption{Results for 1, 2, and 5 trials using only the first block in each
      training fold for training.}
    \end{subfigure}
    \smallskip

    \begin{subfigure}{\linewidth}
      \input{figures/stbf_struct/fpi-1.pgf}
      \caption{Results for 1, 2, and 5 trials using all nine training blocks in
      the training folds.}
    \end{subfigure}
    \caption[Average cross-validated \textsc{stbf-struct} accuracy]{%
      Average cross-validated \textsc{stbf-struct} accuracy using
			one trial per block over all 21 subjects
			relative to the number of iterations used to estimate the Kronecker-Toeplitz structured shrunk
			covariance. Shaded area represents a 95\% confidenced interval obtained
      through 1000 bootstrapping iterations.
      Accuracy does not improve when using more than one iteration.}
		\label{fig:iterations}
	\end{figure}

	\subsection{Classifier accuracy for limited training data}
	It is of interest to keep the calibration time before \ac{bci}
	operation as short as possible.
	We mimic this problem by training the classifier with as few training epochs as possible.
	We evaluate the performance of all classifiers for different levels of
	available training data and apply the cross-validation procedure nine times (the number of blocks in the training fold) for all subjects, keeping the
	corresponding number of blocks in the training folds and dropping the rest.
  \Cref{fig:accuracy}, \cref{fig:stbf-struct/accuracy-ap} and
  \crefrange{tab:stbf-struct/accuracies/stbf-struct}{tab:stbf-struct/accuracies/xdawn-rg}
  shows each classifier's accuracy relative to the data availability and trials
  used for agveraging.
	We statistically compare the two newly proposed classifiers,
	\textsc{stbf-struct} and \textsc{stbf-shrunk} with different levels of training
	data availability using a one-sided Wilcoxon signed-rank test with Holm correction for the multiple pairwise comparisons between classifiers.
	Statistical comparisons were performed for trial
  \cref{tab:stbf-struct/p-values-1}, two \cref{tab:stbf-struct/p-values-2}
  and five trials \cref{tab:stbf-struct/p-values-5}.

  \begin{figure}
    %\includegraphics[width=.66\linewidth]{figures/stbf_struct/accuracy.eps}
    \sffamily
    \input{figures/stbf_struct/accuracy.pgf}
    \caption[Clasifier accuracy in function of available training data.]{%
      Accuracy of the different classifiers for all 21 subjects relative to the
			number of blocks available for training. One block consists of 135
			epochs and corresponds to 27 seconds of stimulation. Accuracies
			are shown for the evaluation settings averaging over 1, 2, and
			3 trials of testing stimuli.
			\cref{fig:stbf-struct/accuracy-ap} contains results for all numbers of trials.
			While \textsc{stbf-emp} is unstable when little training data
			are available,
			regularization of the covariance matrix (\textsc{stbf-shrunk} and
			\textsc{stbf-struct}) drastically improves performance.}
		\label{fig:accuracy}
	\end{figure}

>>>>

	\section{Materials \& methods}
	\subsection{Notation}
	We represent matrices with bold capital letters, vectors with bold
	lowercase letters, fixed scalars with uppercase cursive letters and variable
  scalars with cursive lowercase letters.
	The epoched \ac{eeg} data with $N$ epochs, $C$ channels, and $S$ samples are
  represented in epoch format as $\{\mat{X}_n\in\mathbb{R}^{C\times S}\}^N_{n=1}$
  or flattened vector format by concatenating all channels for each epoch.
	Flattening results in $\{\mat{x}_n\in\mathbb{R}^{CS}\}^N_{n=1}$ such that
  $\mat{x}_n = \vect(\mat{X}_n)$.
	The real covariance matrix of the epochs in vector format is
  denoted by $\mat{C}$, estimators thereof as $\hat{\mat{C}}$.

	\subsection{Spatiotemporal beamforming}
  \ac{lcmv}-beamforming was initially introduced to \ac{eeg} analysis as a filter for
  source localization~\cite{VanVeen1997} to enhance the \ac{snr}.
	\textcite{VanVliet2015} first applied the spatiotemporal
	\ac{lcmv}-beamformer as a method for the analysis of \acp{erp}.
	The extension to the combined spatiotemporal domain~\cite{VanVliet2015} and the
	data-driven approaches proposed by \textcite{Treder2016} and
	\textcite{Wittevrongel2016} allow for its application to classification problems.

	For the following analyses, we assume that all \ac{eeg} channels are normalized with zero mean and unit variance without loss of generality.
	Solving \cref{eq:stbf-struct/minimum-variance} under the linear constraint given by
  \cref{eq:stbf-struct/linear-constraint} returns the filter weights $\mat{w}$ defining the spatiotemporal \ac{lcmv}-beamformer.
	\begin{equation}
    \argmin_\mat{w}\mat{w}^\intercal \mat{C}
		\mat{w}^\intercal
		\label{eq:stbf-struct/minimum-variance}
	\end{equation}
	\begin{equation}
		\mat{a}^\intercal\mat{w} = 1
		\label{eq:stbf-struct/linear-constraint}
	\end{equation}
	These weights minimize the variance of the output of the filter while enhancing
	the signal characterized by the constraint.
  $\mat{a} = \vect(\mat{A})$ is the data-driven activation pattern, a template
	of the signal of interest maximizing the difference between two classes of
  epochs, here denoted as the `target' (T) and `non-target' (NT) class.
  The activation pattern is then determined as follows:
	\begin{equation}
		\mathbf{a} =
    \frac{1}{N_\mathrm{T}}\sum_\mathrm{T}\mathbf{x_n} -
    \frac{1}{N_\mathrm{NT}}\sum_\mathrm{NT}\mathbf{x_n}
		\label{eq:stbf-struct/ap}
	\end{equation}

@@

	\subsubsection{Shrunk covariance estimation}
	\label{sec:stbf-struct/methods/shrunk-cov}
	The shrinkage covariance estimator creates a better conditioned inversion matrix problem and generally performs better when applied to unseen data.
	The estimators for the covariance and inverse covariance are given by:
	\begin{equation}
    \hat{\mat{C}}_\alpha =
    (1-\alpha) \hat{\mat{C}}_\text{emp}
    + \alpha\frac{\Tr(\hat{\mat{C}}_\text{emp})}{CS}\mathbb{I}
		\label{eq:shrinkage}
	\end{equation}
	\begin{equation}
    \widehat{\mat{C}^{-1}}_\alpha =
    \hat{\mat{C}}^+_\alpha
		\label{eq:stbf-struct/shrinkage-inv}
	\end{equation}
	with $0<\alpha<1$.
	Analogous to $L_2$ regularization of the beamforming problem,
	shrinkage reduces the ratio between the smallest and largest eigenvalues
	of the covariance matrix by strengthening the diagonal.
	\cref{fig:kronlda-covs}c and \cref{fig:kronlda-covs}d respectively show examples of the
	shrunk estimator of the covariance and the inverse covariance matrices.

@@

	Methods exist to estimate an optimal shrinkage value from the data directly.
	Most notable among these are the Ledoit-Wolf procedure~\cite{Ledoit2004},
	Rao-Blackwell Ledoit-Wolf~\cite{Chen2010}, and Oracle Approximating Shrinkage~\cite{Chen2010}.
  A more recent estimation method~\cite{Tong2018} emulates a \ac{loocv} scheme expressed by the data-driven closed-form
	estimate:
	\begin{equation}
		  \alpha =
		  1-\frac{%
        \splitfrac{
          \frac{N}{N-1}\Tr(\hat{\mat{C}}_\text{emp}^2)
          - \frac{2}{CS}\left[\Tr(\hat{\mat{C}}_\text{emp})\right]^2
          + \frac{1}{CS}\Tr(\hat{\mat{C}}_\text{emp}^2)
        }{
		  	  - \frac{1}{N(N-1)}\sum_{n=1}^N||\mat{x}_n||_2^4
        }
		  }{%
        \splitfrac{
          \frac{N^2 -2N}{(N-1)^2}\Tr(\hat{\mat{C}}_\text{emp}^2)
          - \frac{2}{CS}\left[\Tr(\hat{\mat{C}}_\text{emp})\right]^2
          + \frac{1}{CS}\Tr(\hat{\mat{C}}_\text{emp}^2)
        }{%
		  	  + \frac{1}{N(N-1)^2}\sum_{n=1}^N||\mat{x}_n||_2^4
        }
		  }
		\label{eq:loocv}
	\end{equation}
	We opt for the \ac{loocv} shrinkage estimator because it avoids some of the
	assumptions made by \textcite{Ledoit2004} and \textcite{Chen2010} and
	because it generalizes to structured covariance estimation as described in
	\cref{sec:stbf-struct/methods/structured-estimation}.

@@

	Combined, this leads to the iterative estimation algorithm described by the
	following equations:
	\begin{subequations}
		\begin{equation}
      \tilde{\mat{S}}_{k+1} =
			\frac{1}{N}
      \sum^N_{n=1}\mat{X}_n^\intercal\hat{\mat{T}}_k^+\mat{X}_n
			\label{eq:stbf-struct/fpi-spatial}
		\end{equation}
		\begin{equation}
      \tilde{\mat{T}}_{k+1} =
			\frac{1}{N}
      \sum^N_{n=1}\mat{X}_n\hat{\mat{S}}_k^+\mat{X}_n^\intercal
			\label{eq:stbf-struct/fpi-temporal}
		\end{equation}
	\end{subequations}
	\begin{subequations}
		\begin{equation}
      \tilde{\mat{S}}_{k+1}^{(\beta)} =
      (1-\beta_{k+1})\tilde{\mat{S}}_{k+1}
      +\beta_{k+1}\frac{\Tr(\tilde{\mat{S}}_{k+1})}{C}\mathbb{I}
			\label{eq:stbf-struct/fpi-spatial-shrunk}
		\end{equation}
		\begin{equation}
      \tilde{\mat{T}}_{k+1}^{(\gamma)} =
      (1-\gamma_{k+1})\tilde{\mat{T}}_{k+1}
      +\gamma_{k+1}\frac{\Tr(\tilde{\mat{T}}_{k+1})}{S}\mathbb{I}
			\label{eq:stbf-struct/fpi-temporal-shrunk}
		\end{equation}
	\end{subequations}
	\begin{subequations}
		\begin{equation}
      \hat{\mat{S}}_{k+1} =
      \frac{C}{\Tr\left[\tilde{\mat{S}}_{k+1}^{(\beta)}\right]}
      \tilde{\mat{S}}_{k+1}^{(\beta)}
			\label{eq:stbf-struct/fpi-spatial-norm}
		\end{equation}
		\begin{equation}
      \hat{\mat{T}}_{k+1} =
      \frac{S}{\Tr\left[\tilde{\mat{T}}_{k+1}^{(\gamma)}\right]}
      \tilde{\mat{T}}_{k+1}^{(\gamma)}
			\label{eq:stbf-struct/fpi-temporal-norm}
		\end{equation}
	\end{subequations}
  $\hat{\mat{S}}_0$ and $\hat{\mat{T}}_0$ can be initialized to any positive definite matrix.
	We choose to use the identity matrices $\mathbb{I}^{C\times C}$ and $\mathbb{I}^{S\times S}$.
  After each iteration, all diagonals of $\hat{\mat{R}}_{k+1}$ are set to their mean
  values to ensure that $\hat{\mat{R}}_{k+1}$ and $\hat{\mat{T}}_{k+1}$ are Toeplitz structured.

@@

	\begin{subequations}
		\begin{equation}
			\beta_{k+1} =
			1-
			\frac{
        \splitfrac{
          \frac{N}{N-1}\Tr(\tilde{\mat{\hat{\S}}}_{k+1}^2)
          - \frac{2}{C}\left[\Tr(\mat{\tilde{S}}_{k+1})\right]^2
          + \frac{1}{C}\Tr(\tilde{\mat{S}}_{k+1}^2)
        }{%
				  - \frac{1}{N(N-1)}\sum_{i=1}^N
          \left[\Tr(\mat{X}_i\hat{\mat{T}}_k^+\mat{X}_i^\intercal)^2\right]
        }
			}{
        \splitfrac{
          \frac{N^2-2N}{(N-1)^2}\Tr(\tilde{\mat{S}}_{k+1}^2)
          - \frac{2}{C}\left[\Tr(\tilde{\mat{S}}_{k+1})\right]^2
          + \frac{1}{C}\Tr(\tilde{\mat{S}}_{k+1}^2)
        }{%
				  + \frac{1}{N(N-1)^2}\sum_{i=1}^N
          \left[\Tr(\mat{X}_i\hat{\mat{T}}_k^+\mat{X}_i^\intercal)^2\right]
        }
			}
			\label{eq:stbf-struct/spatial-shrinkage}
		\end{equation}
		\begin{equation}
			\gamma_{k+1} =
			1-
			\frac{
        \splitfrac{
          \frac{N}{N-1}\Tr(\tilde{\mat{T}}_{k+1}^2)
          - \frac{2}{S}\left[\Tr(\tilde{\mat{T}}_{k+1})\right]^2
          + \frac{1}{S}\Tr(\tilde{\mat{T}}_{k+1}^2)
        }{%
				  - \frac{1}{N(N-1)}\sum_{n=1}^N
          \left[\Tr(\mat{X}_n^\intercal\mat{\mat{S}}_k^+\mat{X}_n)^2\right]
        }
			}{
        \splitfrac{
          \frac{N^2-2N}{(N-1)^2}\Tr(\tilde{\mat{T}}_{k+1}^2)
          - \frac{2}{S}\left[\Tr(\tilde{\mat{T}}_{k+1})\right]^2
          + \frac{1}{S}\Tr(\tilde{\mat{T}}_{k+1}^2)
        }{
				  + \frac{1}{N(N-1)^2}\sum_{n=1}^N
          \left[\Tr(\mat{X}_n^\intercal\hat{\mat{S}}_k^+\mat{X}_n)^2\right]
        }
			}
			\label{eq:stbf-struct/temporal-shrinkage}
		\end{equation}
	\end{subequations}
	The shrinkage parameters $0<\beta_{k+1}<1$ and $0<\gamma_{k+1}<1$ should be
	re-determined after each iteration.
	The Oracle Approximation Shrinkage method can also be used to determine
	$\beta_{k+1}$ and $\gamma_{k+1}$~\cite{Chen2010,Xie2021} but performs worse for spatiotemporal \ac{eeg} data since not all assumptions are met.

@@

  \subsubsection{\Acl{stbf} classifier}
	Before calculating the \ac{stbf}, the signal was downsampled to
	32 Hz or twice the low-pass frequency 16 Hz, resulting in 17 time samples
	between 0.1 and 0.6s. According to the Nyquist Theorem, more samples would not
	contain more information hence the minimum temporal resolution is chosen to reduce
	the dimensionality of the covariance and improve its condition number.
	The activation pattern is the difference between the averages of epochs in response to cued targets and the averages of those in response to non-cued targets.
	We constructed three variations of the spatiotemporal beamformer the \ac{stbf-emp} as in
	\cref{sec:stbf-struct/methods/emp-cov}, the \ac{stbf-shrunk} as in
	\cref{sec:stbf-struct/methods/shrunk-cov}, and the \ac{stbf-struct} with \ac{loocv} shrinkage for
	the Kronecker factors as in \cref{sec:stbf-struct/methods/struct-cov}.

	\subsubsection{Riemannian geometry classifier}
	\label{sec:riemannian}
	We opted for a Riemannian
	geometry-based classifier to compare our results.
	The Riemannian model (XDAWN+RG) uses the XDAWN spatial filter combined
	with Riemannian geometry in tangent space as implemented by Barachant et
	al.~\cite{Barachant2014a}.
	This classifier uses four XDAWN spatial filters and each epoch's empirical spatial covariance matrix.
	The target with the maximum score is the prediction of the cued target.
	XDAWN+RG was trained and validated without downsampling using epochs
	at the original sample rate of 2048 Hz.
  Preliminary experimentation revealed that best performance in low and high
  data availability setting was achieved when no shrinkage was applied to the estimation of XDAWN filters or the subsequent
  covariance matrices for this specific dataset.
  Hence, no shrinkage regularization will be applied to XDAWN+RG in
  further experiments.

	\section{Results}
	\subsection{Minimum required fixed-point iterations}
	The fixed point iteration algorithm described
	in
  \crefrange{eq:stbf-struct/fpi-spatial}{eq:stbf-struct/fpi-temporal-norm}
  estimates the Kronecker-Toeplitz structured covariance for the
	\ac{stbf-struct} classifier.
	Fixed-point iteration is an iterative procedure starting from (in our case)
	non-informed initial guesses for the spatial and temporal covariance matrices.
	As a stopping criterion, one could impose a threshold on the difference in
	outcome of successive steps, e.g., based on the covariance norm or the
	classifier accuracy.
	However, few iterations or even just one~\cite{Castaneda2014} suffice to achieve satisfactory performance in practice.

	\cref{fig:iterations} confirms these results for the \ac{stbf-struct} classifier.
	Using more than one fixed-point iteration does not significantly improve the
	accuracy across the amounts of training data and the number of trials
  used for evaluation.
	Hence, only one iteration is used for the \ac{stbf-struct} classifier, leading to a drastic speed-up of
  the  training process.

  \begin{figure}[t]
    \begin{subfigure}{\linewidth}
      \input{figures/stbf_struct/fpi-0.pgf}
      \caption{Results for 1, 2, and 5 trials using only the first block in each
      training fold for training.}
    \end{subfigure}
    \smallskip

    \begin{subfigure}{\linewidth}
      \input{figures/stbf_struct/fpi-1.pgf}
      \caption{Results for 1, 2, and 5 trials using all nine training blocks in
      the training folds.}
    \end{subfigure}
    \caption[Average cross-validated \acs{stbf-struct} accuracy]{%
      Average cross-validated \ac{stbf-struct} accuracy using
			one trial per block over all 21 subjects
			relative to the number of iterations used to estimate the Kronecker-Toeplitz structured shrunk
			covariance. Shaded area represents a 95\% confidenced interval obtained
      through 1000 bootstrapping iterations.
      Accuracy does not improve when using more than one iteration.}
		\label{fig:iterations}
	\end{figure}

	\subsection{Classifier accuracy for limited training data}
	It is of interest to keep the calibration time before \ac{bci}
	operation as short as possible.
	We mimic this problem by training the classifier with as few training epochs as possible.
	We evaluate the performance of all classifiers for different levels of
	available training data and apply the cross-validation procedure nine times (the number of blocks in the training fold) for all subjects, keeping the
	corresponding number of blocks in the training folds and dropping the rest.
  \Cref{fig:accuracy}, \cref{fig:stbf-struct/accuracy-ap} and
  \crefrange{tab:stbf-struct/accuracies/stbf-struct}{tab:stbf-struct/accuracies/xdawn-rg}
  shows each classifier's accuracy relative to the data availability and trials
  used for agveraging.
	We statistically compare the two newly proposed classifiers,
	\ac{stbf-struct} and \ac{stbf-shrunk} with different levels of training
	data availability using a one-sided Wilcoxon signed-rank test with Holm correction for the multiple pairwise comparisons between classifiers.
	Statistical comparisons were performed for trial
  \cref{tab:stbf-struct/p-values-1}, two \cref{tab:stbf-struct/p-values-2}
  and five trials \cref{tab:stbf-struct/p-values-5}.

  \begin{figure}
    %\includegraphics[width=.66\linewidth]{figures/stbf_struct/accuracy.eps}
    \sffamily
    \input{figures/stbf_struct/accuracy.pgf}
    \caption[Clasifier accuracy in function of available training data.]{%
      Accuracy of the different classifiers for all 21 subjects relative to the
			number of blocks available for training. One block consists of 135
			epochs and corresponds to 27 seconds of stimulation. Accuracies
			are shown for the evaluation settings averaging over 1, 2, and
			3 trials of testing stimuli.
			\cref{fig:stbf-struct/accuracy-ap} contains results for all numbers of trials.
			While \ac{stbf-emp} is unstable when little training data
			are available,
			regularization of the covariance matrix (\ac{stbf-shrunk} and
			\ac{stbf-struct}) drastically improves performance.}
		\label{fig:accuracy}
	\end{figure}

	The tables show that \ac{stbf-struct} has a significant advantage over
  \ac{stbf-shrunk} when the number of training blocks is low, with
  significance $p=0.005$ and  an accuracy increase of 4,14$\%.$ for a single training block
  and testing trial.
  This effect is present for 1-, 2- and 5-trial evaluation, up to 3, 4 and 3
  blocks respectively, but the advantage decreases when adding more training blocks.
	Both \ac{stbf-struct} and \ac{stbf-shrunk} perform significantly better
  than \ac{stbf-emp} for all evaluated settings, with $p<0.001$ for all number
  of training blocks and testing trials, with accuracy
  increase of 6.92 and 2.78$\%.$ respectively for a single training block
  and testing trial.
	Compared to XDAWN+RG, \ac{stbf-struct} also has significantly
	higher accuracy in almost all evaluated settings, except when using only one
  training block, with significance $p<0.001$ and an accuracy increase of
  5.20$\%.$ when using all 9 training blocks and 1 testing trial.
	\ac{stbf-shrunk} does not outperform XDAWN+RG when training data
  is low (up to 3,3 and 2 blocks respectively for 1, 2 and 5 testing trials),
  but gains a significant advantage	when using more training data, with
  significange $p=0.001$ and an accuracy increase of 5.33$\%.$ when using all
  9 training blocks and 1 testing trial.
